---
title: "Iteration"
output: html_notebook
---

```{r setup}
library(tidyverse)

# Toy data
set.seed(1000)
exams <- list(
  student1 = round(runif(10, 50, 100)),
  student2 = round(runif(10, 50, 100)),
  student3 = round(runif(10, 50, 100)),
  student4 = round(runif(10, 50, 100)),
  student5 = round(runif(10, 50, 100))
)

exams.df <- data_frame(
  student1 = round(runif(10, 50, 100)),
  student2 = round(runif(10, 50, 100)),
  student3 = round(runif(10, 50, 100)),
  student4 = round(runif(10, 50, 100)),
  student5 = round(runif(10, 50, 100))
)

extra_credit <- list(0, 0, 10, 10, 15)
```

## Your Turn 1

Here is a list:

```{r}
a_list <- list(num = c(8, 9), 
            log = TRUE,    
            cha = c("a", "b", "c"))
```

Here are two subsetting commands. Do they return the same values? Run the code chunk above, _and then_ run the code chunks below to confirm

```{r}
a_list["num"]
```

```{r}
a_list$num
```

## Your Turn 2

What will each of these return? Run the code chunks to confirm.

```{r}
vec <- c(-2, -1, 0, 1, 2)
abs(vec)
```

```{r}
lst <- list(-2, -1, 0, 1, 2)
abs(lst)
```

## Your Turn 3

Run the code in the chunks. What does it return?

```{r}
list(student1 = mean(exams$student1),
     student2 = mean(exams$student2),
     student3 = mean(exams$student3),
     student4 = mean(exams$student4),
     student5 = mean(exams$student5))
```

```{r}
map(exams, mean)
```

## Your Turn 4

Calculate the variance (`var()`) of each student’s exam grades.

```{r}
exams %>% 
   map(var)
```


```{r}
exams %>% 
  map_dbl(mean)

exams %>% 
  map_dbl(quantile, prob=.8)

exams %>% 
  map(length) %>% 
  map_lgl(all.equal,10)
```








## Your Turn 5

Calculate the max grade (max())for each student. Return the result as a vector.

```{r}
exams %>% 
  map_dbl(max)
```


```{r}
vec <- exams$student1
(sum(vec) - min(vec))/(length(vec) - 1)

vec <- exams$student1
grade1 <- function(){
  (sum(vec) - min(vec))/(length(vec) - 1)
}

grade2 <- function(vec){
  (sum(vec) - min(vec))/(length(vec) - 1)
}

grade3 <- function(x){
  (sum(x) - min(x))/(length(x) - 1)
}

```





## Your Turn 6

Write a function that counts the best exam twice and then takes the average. Use it to grade all of the students.

1. Write code that solves the problem for a real object  
2. Wrap the code in `function(){}` to save it  
3. Add the name of the real object as the function argument 

```{r}
nums <- exams$student4
(sum(nums) + max(nums))/(length(nums)+1)

best <- function(){
  (sum(nums) + max(nums))/(length(nums)+1)
}

best <- function(nums){
  (sum(nums) + max(nums))/(length(nums)+1)
}

best <- function(x){
  (sum(x) + max(x))/(length(x)+1)
}

exams %>% 
  map(best)

exams %>% 
  map_dbl(best)
```


### Extra work

Drop 3 minimum scores and then find the average 


Step 1: solve the problem for a real object

```{r}
vec <- exams$student1
vec.3min <- sort(vec)[1:3]
(sum(vec) - sum(vec.3min))/(length(vec)-length(vec.3min))
```

Step 2: Wrap the code in a function call to save it 

```{r}
vec <- exams$student1
drop3min <- function(){
  vec.3min <- sort(vec)[1:3]
(sum(vec) - sum(vec.3min))/(length(vec)-length(vec.3min))
}
```

Step 3: Add the name of the real object as function argument

```{r}
vec <- exams$student1
drop3min <- function(vec){
  vec.3min <- sort(vec)[1:3]
(sum(vec) - sum(vec.3min))/(length(vec)-length(vec.3min))
}
```

Step 4: Replace arguments with x,y,z ......

```{r}
drop3min <- function(x){
  x.3min <- sort(x)[1:3]
(sum(x) - sum(x.3min))/(length(x)-length(x.3min))
}
```


Apply the function


```{r}
exams %>% 
  map_dbl(drop3min)
```

### Make the above function more robust so that it can deal with missing values

```{r}
exams$student3[6] <- NA
```

```{r}
exams %>% 
  map_dbl(drop3min)
```

Step 1: solve the problem for a real object

```{r}
vec <- exams$student3
vec[is.na(vec)] <- 0
vec.3min <- sort(vec)[1:3]
(sum(vec) - sum(vec.3min))/(length(vec)-length(vec.3min))
```

Step 2: Wrap the code in a function call to save it 

```{r}
vec <- exams$student3
drop3min.rob <- function(){
  vec[is.na(vec)] <- 0
  vec.3min <- sort(vec)[1:3]
(sum(vec) - sum(vec.3min))/(length(vec)-length(vec.3min))
}
```

Step 3: Add the name of the real object as function argument

```{r}
vec <- exams$student1
drop3min.rob <- function(vec){
  vec[is.na(vec)] <- 0
  vec.3min <- sort(vec)[1:3]
(sum(vec) - sum(vec.3min))/(length(vec)-length(vec.3min))
}
```

Step 4: Replace arguments with x,y,z ......

```{r}
drop3min.rob <- function(x){
  x[is.na(x)] <- 0
  x.3min <- sort(x)[1:3]
(sum(x) - sum(x.3min))/(length(x)-length(x.3min))
}
```


Apply the function


```{r}
exams %>% 
  map_dbl(drop3min.rob)
```

# Modify the function so that the number of scores to drop is also inbuilt in the function


Step 1

```{r}
vec <- exams$student1
num <- 5
vec[is.na(vec)] <- 0
vec.drop <- sort(vec)[1:num]
(sum(vec) - sum(vec.drop))/(length(vec)-length(vec.drop))
```

Step 2

```{r}
drop.custom <- function(){
vec[is.na(vec)] <- 0
vec.drop <- sort(vec)[1:num]
(sum(vec) - sum(vec.drop))/(length(vec)-length(vec.drop))
}
```

Step 3

```{r}
drop.custom <- function(vec,num){
vec[is.na(vec)] <- 0
vec.drop <- sort(vec)[1:num]
(sum(vec) - sum(vec.drop))/(length(vec)-length(vec.drop))
}
```


Step 4

```{r}
drop.custom <- function(x,y){
x[is.na(x)] <- 0
x.drop <- sort(x)[1:y]
(sum(x) - sum(x.drop))/(length(x)-length(x.drop))
}
```

Apply the function 


```{r}
exams %>% 
  map_dbl(drop.custom,4)
```




### Your Turn 7

Compute a final grade for each student, where the final grade is the average test score plus any `extra_credit` assigned to the student. Return the results as a double (i.e. numeric) vector.

```{r}

add_extra <- function(x,y){
  drop3min(x) + y
} 

exams.df %>% 
  map2_dbl(extra_credit, add_extra)
```


***

# Take Aways

Lists are a useful way to organize data, but you need to arrange manually for functions to iterate over the elements of a list.

You can do this with the `map()` family of functions in the purrr package.

To write a function, 

1. Write code that solves the problem for a real object  
2. Wrap the code in `function(){}` to save it  
3. Add the name of the real object as the function argument 

This sequence will help prevent bugs in your code (and reduce the time you spend correcting bugs). 